Overlapping Batches for the Assessment of Solution Quality in Stochastic Programs

### 2014-07-01 Referee2_Response.tex

\begin{itemize}
\item[] \textit{4, 51: ``discuss overlapping''}
\end{itemize}

\noi
Done.  
\medskip

\begin{itemize}
\item[] \textit{4, 54 (and elsewhere): When you refer to ``variable amounts of overlap'', that is really nothing new. Certainly, any overlapping estimator will have end effects in which the overlap amounts differ from those in the ``middle'' region of the data.}
\end{itemize}

\noi
%You are right.
We meant to say that the degree of overlap is parametrized and this parameter can be chosen by the user. 
We reworded this sentence to reflect this. 
xxx  We also \ldots avoided using this term and replaced by. . . 
\medskip

@@

\begin{itemize}
\item[] \textit{4, 57 (and elsewhere): ``classical case'' refers to nonoverlapping batch means, right?}
\end{itemize}

\noi
No; by `classical case', we meant the use of overlapping batches in simulation output analysis (as opposed to the assessment of solution quality in stochastic programming). 
We reworded this sentence to make it clear. 
Please note that based on your earlier comments, we reworded this sentence to avoid using the term `variance reduction'. 
We also went through the paper and removed several uses of the word `classical' that referred to nonoverlapping batches to reduce ambiguity.
%(as these were referring to nonoverlapping batches). 
%In the current version of the paper, `classical case' is used to only indicate  the use of overlapping batches in simulation output analysis. 
\medskip

\begin{itemize}
\item[] \textit{5, 12: Does $\sigma^2$ refer to the marginal variance or the ``variance parameter'' (sum of the covariances at all lags)?}
\end{itemize}

\noi
Variance parameter. 
We revised this sentence and added a footnote defining the variance parameter for readers who might not be familiar with this term (e.g., readers with more of an optimization background). 
\medskip

### 2014-07-01 omrp_lb_rev.tex
Major Comment 2 of Referee 2
Removed Section 3.1 and summarized everything in Section 2.1.  Also,
removed some references/text from the later sections to Sec 2.1, as we
use them later on. . .  This way, we have a self-contained background on
overlapping batches that lists the main results we use in the paper.

% Commands for probability
\newcommand{\p}[1]{\mathbb{P} \left\{ #1 \right\}}
\newcommand{\e}[1]{\mathbb{E} \left[ #1 \right]
}
\newcommand{\ee}[2]{\mathbb{E}_{#1} \left[ #2 \right]}
\newcommand{\var}[1]{\mathrm{Var} \left( #1 \right)}
\newcommand{\cov}[1]{\mathrm{Cov} \left( #1 \right)}
\newcommand{\cor}[1]{\mathrm{Corr} \left( #1 \right)}
\newcommand{\varh}[1]{\widehat{\mathrm{Var}} \left( #1 \right)}
\newcommand{\vart}[1]{\widetilde{\mathrm{Var}} \left( #1 \right)}
\newcommand{\vartg}[1]{\widetilde{\mathrm{Var}}_\gamma \left( #1 \right)}
\newcommand{\varhat}{\widehat{\mathrm{Var}}}

@@

\subsection{Overlapping Batch Means} \label{ssec:obm}
Consider a covariance stationary stochastic process with mean $\mu$ and variance $\sigma^2$.  
The task is to estimate $\mu$ given some realization of the univariate stochastic process $\y = (y^1, y^2, \dots, y^n)$.  
Typically, this involves forming a CI of the form $[L_\alpha(\y), U_\alpha(\y)]$ for a given level of significance $\alpha$ such that $\p{L_\alpha(\y) \leq \mu \leq U_\alpha(\y)} = 1 - \alpha$ or that this probability is approximately $1 - \alpha$ for large enough sample sizes.  
The usual estimator for $\mu$ is the sample mean $\ybb = \frac{1}{n} \sum_{i=1}^n y^i$, which is an unbiased estimator even in the presence of correlated data.  
However, the usual variance estimator $\frac{1}{n-1} \sum_{i=1}^n (y^i - \ybb)^2$ can be inappropriate to use in the presence of correlated data, resulting in inaccurate interval estimators.  
To overcome this difficulty, various variance estimators have been proposed in the simulation literature; see, e.g., \citep{law_07}.  
We briefly review two of these variance estimators: nonoverlapping and overlapping batch means.

Nonoverlapping batch means takes $n$ observations, $y^1, \dots, y^n$ of the stochastic process and splits them into $k$ batches of size $m$, where $k = \frac{n}{m}$ (for simplicity, assume, for now, $n = mk$).  
The sample mean of each batch is then computed as
$
	\yb_j = \frac{1}{m} \sum_{i=1}^{m} y^{m(j-1)+i}, j = 1,2, \dots, k,
$
and the overall sample mean $\ybb = \frac{1}{k} \sum_{j=1}^k \yb_j = \frac{1}{n} \sum_{i=1}^n y^i$ provides a point estimator of $\mu$.  
The variance estimator of the batch means, which is $m/n$ times the sample variance, is calculated as
\begin{equation} \label{eq:var}
	\varhat(\ybb) = \frac{m}{n}\frac{1}{k-1} \sum_{j=1}^k \left( \yb_j - \ybb \right)^2.
\end{equation}
The $(1-\alpha)$-level approximate CI is then formed by $\ybb \pm t_{k-1,\alpha/2} \sqrt{\varhat(\ybb)}$, where $t_{k-1,\alpha/2}$ denotes the $1-\alpha/2$ quantile of the Student's t distribution with $k-1$ degrees of freedom.  
Under appropriate conditions, $n\varhat(\ybb)$ is a consistent estimator of $\sigma^2$.

Overlapping batch means modifies this idea by taking batches given by $\yb_j(m) = \frac{1}{m} \sum_{i=1}^{m}$ $y^{(j-1)+i}$, for $j = 1, \dots, n-m+1$.
As an example, suppose each batch contains $m$ observations and that there are a total of $n=2m$ observations.
In the nonoverlapping case, there are only two batches: the first batch contains observations $y^1, y^2, \dots, y^m$ and the second batch contains observations $y^{m+1}, y^{m+2}, \dots, y^{2m}$.   
In the overlapping batches of \citet{Meketon1984}, the first batch is as before, the second batch now consists of $y^2, y^3, \dots, y^{m+1}$, the third batch contains observations $y^3, y^4, \dots, y^{m+2}$, and so on.  
This is what we call the \emph{maximally-overlapping} batch means method, because $m-1$ observations are common to---and only one observation changes between---adjacent batches.

The variance estimator is updated to
\begin{equation} \label{eq:overlap_variance_def}
	\vart{\ybb} = \frac{1}{(\tfrac{n}{m}-1)(n - m + 1)}\sum_{j=1}^{n-m+1} (\yb_j(m) - \ybb)^2.
\end{equation}
The corresponding overlapping batch means are no longer independent; however, by simply reusing the data in this fashion---and seemingly in a contradictory way that induces dependence---one obtains a better variance estimator. 
\citet{Meketon1984} show that: (i) the overlapping variance estimator has nearly the same bias as the standard nonoverlapping variance estimator, and (ii) the overlapping estimator has only two-thirds of the asymptotic variance of the nonoverlapping one.  
That is, 
\begin{equation} \label{eq:asym_var}
	\frac{ \var{\vart{\ybb}} }{ \var{\varhat(\ybb)} } \rightarrow \frac{2}{3}
\end{equation}
in the limit as batch size $m$ and the number of batches $(n/m)$ tend to infinity.  
Note that the degrees of freedom in (\ref{eq:overlap_variance_def}) is slightly different than in \citep{Meketon1984}.  
The degrees of freedom in (\ref{eq:overlap_variance_def}) make $\vart{\ybb}$ an unbiased estimator for i.i.d.\ data for all finite $m$ and $n$ with the same asymptotic benefits \citep{Song1992}.  
The $(1-\alpha)$-level approximate CI is similarly formed by $\ybb \pm t_{3(k-1)/2,\alpha/2} \sqrt{\vart{\ybb}}$, with an adjustment to the degrees of freedom.

\citet{Welch1987} established the relationship between OBM and spectral estimators and also considered variable amounts of overlap, see also \citep{Song1992}.  
%In our application to assessment of solution quality, we also consider varying the amount of overlap between neighboring batches.  
We defer this discussion on variable overlap to \S \ref{ssec:overlap} and continue with a brief review of MRP.

@@

Taking $n$ observations of the random variables $\xit$ and splitting them into $k$ nonoverlapping batches of size $m$, define 
\begin{align} 
	\gb_j  & = \frac{1}{m} \sum_{i = 1}^{m} f(\xh, \xit^{m(j-1)+i}) - \min_{\x \in X} \frac{1}{m} \sum_{i=1}^{m} f(\x, \xit^{m(j-1)+i}), \ \ \ \ j = 1,2, \dots, k \label{eq:mrp_gap}\\
         & = \frac{1}{m} \sum_{i = 1}^{m} \left[ f(\xh, \xit^{m(j-1)+i}) -  f(\xs_j, \xit^{m(j-1)+i}) \right], \ \ \ \ j = 1,2, \dots, k, \notag
\end{align}
where $\xs_j$ denotes an optimal solution to (\ref{eq:sto_prog_m}) of batch $j$ given on the right-hand side of (\ref{eq:mrp_gap}).    
Since we are assessing the quality of a given solution $\xh \in \X$, we suppress it from the notation $\gb_j$.
The optimality gap estimator $\gb_j$ is similar to $\yb_j$ with individual observations $y^{m(j-1)+i} = f(\xh,\xit^{m(j-1)+i}) - f(\xs_j,\xit^{m(j-1)+i})$, for $i=1,\ldots,m$ and $j=1,\ldots, k$.
Notice that in this case, $y^i$ not only depends on observation $i$ of batch $j$, $\xit^{m(j-1)+i}$, but all the observations in that batch $\xit^{m(j-1)+1},\xit^{m(j-1)+2},\ldots,\xit^{m(j-1)+m}$ because $\xs_j$ is the optimal solution to (\ref{eq:sto_prog_m}) given in (\ref{eq:mrp_gap}).   
As before, after $k$ estimates are obtained, the overall mean of these estimates $\gbb = \frac{1}{k} \sum_{j=1}^k \gb_j$ provides a point estimator of the optimality gap.  
The variance estimator is obtained as in (\ref{eq:var}) by $\varh{\gbb} = \frac{1}{k} \frac{1}{k-1} \sum_{j=1}^k (\gb_j - \gbb)^2$, which results in an approximate $(1-\alpha)$-level one-sided CI $\left[0, \gbb + t_{k-1,\alpha} \sqrt{\varh{\gbb}} \right]$.

@@

We begin our discussion with variably overlapping batches and then define the estimators of OMRP.

\subsection{Variably Overlapping Batches} 
\label{ssec:overlap}

As before, let $m$ denote the batch size, $n$ the total sample size and $k = \lfloor \tfrac{n}{m} \rfloor$ be the number of nonoverlapping batches.  
In this paper, we use the {\it batch nonoverlap parameter} $1 \leq \gamma \leq m$ to denote how much neighboring batches do {\it not} overlap.  
For instance, $\gamma = m$ corresponds to the case of nonoverlapping batches and $\gamma = 1$ corresponds to the maximally overlapping case of \citet{Meketon1984}.  
In general, $\gamma$ can take any integer value between and including $[1,m]$. 
Let $\lfloor \cdot \rfloor$ be the greatest integer that is less than or equal to its argument. 
The sample mean of each batch estimator is calculated similarly, 
$
\yb_j = \frac{1}{m} \sum_{i=1}^m y^{\gamma(j-1) + i},\ \  j = 1, 2, \dots, \nb, 
$
where $\nb$ is the number of batches used given $n$, $m$, and $\gamma$.  
The variance estimator given in (\ref{eq:overlap_variance_def}) is changed to
\begin{equation} \label{eq:variance_gamma}
	\vartg{\ybb} = \frac{1}{\left( \tfrac{n}{m} - 1 \right) \left( \nb \right)}  \sum_{j=1}^{\nb} (\yb_j - \ybb)^2.
\end{equation}
Note that with $\gamma=1$, (\ref{eq:variance_gamma}) reduces to (\ref{eq:overlap_variance_def}) and with $\gamma=m$ and $n=mk$, (\ref{eq:variance_gamma}) reduces to (\ref{eq:var}).

The amount of asymptotic variance reduction in this estimator depends on the asymptotic ratio of $\gamma/m$, which we denote by $\gammab$.  
For example, when $\gamma = 1$ (or $\gammab=0$; the maximally overlapping case), the variance is reduced to two-thirds ($66.67\%$) of the original nonoverlapping case, as given in (\ref{eq:asym_var}).  
When 75\% of observations overlap (i.e., when $\gamma = m/4$ or $\gammab = 1/4$), the variance is $33/48$th of the original ($68.75\%$), which is near-optimal.   
When only half of the observations overlap ($\gammab = 1/2$) the variance is 75\% of original.  
Let $\lceil \cdot \rceil$ be the smallest integer that is greater than or equal to its argument. 
In general, from the spectral analysis given in \citep{Welch1987}, the variance is reduced to
\begin{equation} \label{eq:var_reduct_formula}
	\gammab \left( 1 + 2 \sum_{j=1}^{\cogb-1} (1-j\gammab)^2 \right) %= -\gammab\left( 1 - 2\gammab\cogb + \gammab^2\cogb^2 \right) + 2\gammab\cogb\left(1-\gammab\cogb\right) + \frac{2}{3}\gammab^3\cogb^3 + \frac{1}{3}\gammab^3\cogb
\end{equation}
of the nonoverlapping case.  
If $\ogb$ is an integer, (\ref{eq:var_reduct_formula}) simplifies to $\frac{2+\gammab^2}{3}$.

A $(1-\alpha)$-level approximate CI on the mean can be formed using variably overlapping batches by $\ybb \pm t_{d_{\gamma}(k-1),\alpha/2}$ $\sqrt{\vartg{\ybb}}$, where the degrees of freedom increase $d_{\gamma}$ is the multiplicative inverse of (\ref{eq:var_reduct_formula}) \citep{Welch1987}.  
If $\ogb$ is an integer, $d_{\gamma} = \frac{3}{2+\gammab^2}$.  For maximally overlapping batches, the degrees of freedom is $d_{\gamma} = \frac{3}{2}$.

\subsection{Definition of Estimators for OMRP}

@@

The results on overlapping batch means occur in the limit as $n, m, k=n/m \rightarrow \infty$ \citep{damerdji1994strong,damerdji1995mean,Meketon1984,Song1992,Welch1987}.  
Let $n_l, m_l$, and $k_l$ be sequences of numbers satisfying these requirements, then the limits will be taken as $l \rightarrow \infty$.  
We assume a batching structure where $m_l = \lfloor n_l ^r \rfloor$ for some $0<r<1$, as is typical in simulation, and $n_l$ tends to infinity as $l \rightarrow \infty$.  
In addition, we may desire that the batch nonoverlap parameter change with $l$, to ensure that $\gamma_l / m_l = \gammab_l$ converges to the constant $\gammab$.  
Now we are ready to define the estimators for OMRP.
\begin{align}
	\gb_{lj} & = \frac{1}{m_l} \sum_{i=1}^{m_l} \left[ f(\xh,\xit^{\gamma_l(j-1)+i}) - f(\xs_j,\xit^{\gamma_l(j-1)+1}) \right],\ \ \ j = 1, 2, \dots, \nbl, \label{eq:gbar} \\
	\gbb_l & = \frac{1}{n_l} \sum_{i=1}^{n_l} \frac{1}{|B(i)|} \sum_{j \in B(i)} \left[ f(\xh,\xiti) - f(\xs_j,\xiti) \right], \label{eq:gbb} \\
	VG_l & = \dfrac{1}{\left( \tfrac{n_l}{m_l} - 1 \right) \left(\nbl\right)} \sum_{j=1}^{\nbl} (\gb_{lj} - \gbb_l)^2. \label{eq:vg}
\end{align}

@@

\subsection{Variance Reduction}
\label{ssec:var_reduct}

\citet{Meketon1984} show that $\frac{n_l}{m_l}\var{n_l VD_l} \rightarrow \alpha \sigma^4_{\xh}$ where $\alpha = \frac{4}{3}$ for maximally overlapping batches and $\alpha = 2$ for nonoverlapping batches, resulting in the decrease of variance for the maximally overlapping batches estimator.  
\citet{Welch1987} expands this result to variable amounts of overlap, showing that $\alpha$ is twice the value of (\ref{eq:var_reduct_formula}).  
\citet{damerdji1995mean} shows that convergence of $\frac{n_l}{m_l}\var{n_l VD_l}$ to the appropriate $\alpha \sigma^4_{\xh}$---thus the mean-square consistency of $n_l VD_l$---is guaranteed for $\tfrac{1}{2} + \tfrac{1}{4+\epsilon} < r < 1$ for both nonoverlapping and maximally overlapping batches under the batching structure $m_l = \lfloor n_l^r \rfloor$ and $4+\epsilon$ finite moments assumption A2$^\prime$.  
(With A2, the range of $r$ that ensures mean-square consistency can be found by setting $\epsilon=0$.)

To prove the variance reduction results in our setting, we wish to show that 
$$
(\tfrac{n_l}{m_l}) \left( \var{n_l VG_l} - \var{n_l VD_l} \right) \rightarrow 0,
$$
for which it suffices to show $\sqrt{{n_l}/{m_l}} \left( n_l VG_l - n_l VD_l \right) \xrightarrow{L^2} 0$, where $\xrightarrow{L^2}$ denotes mean-square convergence.

%To prove the variance reduction results in our setting, we wish to show that $\frac{\frac{n_l}{m_l} \var{n_lVG_l}}{\frac{n_l}{m_l} \var{n_lVD_l}} \rightarrow 1$.  We do this by showing that $({n_l}/{m_l}) \left( \var{n_l VG_l} - \var{n_l VD_l} \right) \rightarrow 0$, for which it suffices to show $\sqrt{{n_l}/{m_l}} \left( n_l VG_l - n_l VD_l \right) \xrightarrow{L^2} 0$, where $\xrightarrow{L^2}$ denotes mean-square convergence.
%
%\begin{theorem} \label{thm:varvar_conv}
%	Suppose assumptions A1--A3 hold, $\gammab = \frac{\gamma_l}{m_l}$ is constant for all $l$, and the batching structure ensures $nVD_l$ converges in mean-square to $\sigma^2_\xh$.  Then $\dfrac{\frac{n_l}{m_l}\var{nVG_l}}{\frac{n_l}{m_l}\var{nVD_l}} \rightarrow 1$ as $l \rightarrow \infty$.
%\end{theorem}

\begin{theorem} \label{thm:varvar_conv}
	Suppose assumptions A1--A3 hold, $\gammab = \frac{\gamma_l}{m_l}$ is constant for all $l$, and the batching structure ensures  $\frac{n_l}{m_l}\var{n_l VD_l} \rightarrow \alpha \sigma^4_{\xh}$ where $\alpha$ is the twice the value of (\ref{eq:var_reduct_formula}). 
        Then, $\frac{n_l}{m_l}\var{n_l VG_l} \rightarrow \alpha \sigma^4_{\xh}$ as $l \rightarrow \infty$ with the same $\alpha$. 
\end{theorem}

@@

Theorem \ref{thm:varvar_conv} gives conditions under which the same asymptotic variance reduction results are achieved for the OMRP optimality gap variance estimator. 
We will empirically examine the variance reduction for small sample sizes in \S \ref{sec:comp}. 
%We end our theoretical properties in the next section by examining the CIs formed using OMRP.

>>>>

% Commands for probability
\newcommand{\p}[1]{\mathbb{P} \left\{ #1 \right\}}
\newcommand{\e}[1]{\mathbb{E} \left[ #1 \right]
}
\newcommand{\ee}[2]{\mathbb{E}_{#1} \left[ #2 \right]}
\newcommand{\var}[1]{\mathrm{Var} \left( #1 \right)}
\newcommand{\cov}[1]{\mathrm{Cov} \left( #1 \right)}
\newcommand{\cor}[1]{\mathrm{Corr} \left( #1 \right)}
\newcommand{\varh}[1]{\widehat{\mathrm{Var}} \left( #1 \right)}
\newcommand{\vart}[1]{\widetilde{\mathrm{Var}} \left( #1 \right)}
\newcommand{\vartg}[1]{\widetilde{\mathrm{Var}}_\gamma \left( #1 \right)}
\newcommand{\vto}[1]{\widetilde{\mathrm{Var}}_{1} \left( #1 \right)}
\newcommand{\vtm}[1]{\widetilde{\mathrm{Var}}_{m} \left( #1 \right)}
\newcommand{\varhat}{\widehat{\mathrm{Var}}}

@@

\subsection{Batch Means} 
\label{ssec:obm}

Consider a covariance stationary stochastic process with mean $\mu$ and variance parameter\footnote{For a given stochastic process $y^1, y^2, \dots$, the variance parameter is the sum of covariances at all lags, $\sigma^2=\mathrm{Var} \left(y^1 \right)+2\sum_{j=1}^{\infty}\mathrm{Cov} \left(y^1,y^{1+j}\right)$. If the variance parameter exists, then  $\sigma^2=\lim_{n\rightarrow \infty}n\mathrm{Var} \left(\ybb\right)$, where $\ybb=\frac{1}{n}\sum_{i=1}^{n} y^{i}$.} $0<\sigma^2<\infty$.  
The task is to estimate $\mu$ given some realization of the univariate stochastic process $\y = (y^1, y^2, \dots, y^n)$.  
Typically, this involves forming a CI of the form $[L_\alpha(\y), U_\alpha(\y)]$ for a given level of significance $\alpha$ such that $\p{L_\alpha(\y) \leq \mu \leq U_\alpha(\y)} = 1 - \alpha$ or that this probability is approximately $1 - \alpha$ for large enough sample sizes.  
The usual estimator for $\mu$ is the sample mean $\ybb = \frac{1}{n} \sum_{i=1}^n y^i$, which is an unbiased estimator even in the presence of correlated data.  
However, the usual variance estimator $\frac{1}{n-1} \sum_{i=1}^n (y^i - \ybb)^2$ can be inappropriate to use in the presence of correlated data, resulting in inaccurate interval estimators.  
To overcome this difficulty, various variance estimators have been proposed in the simulation literature; see, e.g., \citep{law_07}.  
We briefly review two of these variance estimators: nonoverlapping and overlapping batch means.

\subsubsection{Nonoverlapping Batch Means}

Nonoverlapping batch means takes $n$ observations $y^1, \dots, y^n$ of the stochastic process and splits them into $k$ batches of size $m$, where $k = \frac{n}{m}$ (for simplicity, assume, for now, $n = mk$).  
The sample mean of each batch is then computed as
$
	\yb_j = \frac{1}{m} \sum_{i=1}^{m} y^{m(j-1)+i}, j = 1,2, \dots, k,
$
and the overall sample mean $\ybb = \frac{1}{k} \sum_{j=1}^k \yb_j = \frac{1}{n} \sum_{i=1}^n y^i$ provides a point estimator of $\mu$.  
The variance estimator of the batch means, which is $m/n$ times the sample variance, is calculated as
\begin{equation} \label{eq:var}
	\varhat(\ybb) = \frac{m}{n}\frac{1}{k-1} \sum_{j=1}^k \left( \yb_j - \ybb \right)^2.
\end{equation}
The $(1-\alpha)$-level approximate CI is then formed by $\ybb \pm t_{k-1,\alpha/2} \sqrt{\varhat(\ybb)}$, where $t_{k-1,\alpha/2}$ denotes the $1-\alpha/2$ quantile of the Student's t distribution with $k-1$ degrees of freedom.  
Under appropriate conditions, $n\varhat(\ybb)$ is a consistent estimator of $\sigma^2$.

\subsubsection{Overlapping Batch Means}
\label{ssec:overlap}

Overlapping batch means reuses a subset of the observations from one batch to form the next batch. 
As an example, suppose each batch contains $m$ observations and that there are a total of $n=2m$ observations.
In the nonoverlapping case, there are only two batches:
% the first batch contains observations $y^1, y^2, \dots, y^m$ and the second batch contains observations $y^{m+1}, y^{m+2}, \dots, y^{2m}$.  
$y^1, y^2, \dots, y^m$ and $y^{m+1}, y^{m+2}, \dots, y^{2m}$.  
In the overlapping batches of \citet{Meketon1984}, the first batch is as before, the second batch now consists of $y^2, y^3, \dots, y^{m+1}$, the third batch contains observations $y^3, y^4, \dots, y^{m+2}$, and so on.  
This is what we call the \emph{maximally-overlapping} batch means method, because $m-1$ observations are common to---and only one observation changes between---adjacent batches.  
It is possible to vary the amount of overlap between neighboring batches. 
This generalization was first studied by \citet{Welch1987}, who also established the relationship between OBM and spectral estimators; see also \citep{Song1992}.
The overlapping batch means 
%formed using the batch nonoverlap parameter $1 \leq \gamma < m$ 
are no longer independent. 
However, by simply reusing the data in this fashion---and seemingly in a contradictory way that induces dependence---one obtains a better variance estimator. 
That is, the overlapping variance estimator has (i) nearly the same bias as the standard nonoverlapping variance estimator, and it has (ii)
%the overlapping estimator has 
a lower asymptotic variance than its nonoverlapping counterpart.

In this paper, we use the {\it batch nonoverlap parameter} $1 \leq \gamma \leq m$ to denote how much neighboring batches do {\it not} overlap.  
For instance, $\gamma = m$ corresponds to the case of nonoverlapping batches and $\gamma = 1$ corresponds to the maximally-overlapping case of \citet{Meketon1984}.  
In general, $\gamma$ can take any integer value between and including $[1,m]$. 
Let $\lfloor \cdot \rfloor$ be the greatest integer that is less than or equal to its argument. 
The sample mean of each batch is calculated similarly, 
$
\yb_j = \frac{1}{m} \sum_{i=1}^m y^{\gamma(j-1) + i},\ \  j = 1, 2, \dots, \nb, 
$
now taking into account $\gamma$.  
Here, $\nb$ batches are formed given $n$, $m$, and $\gamma$.  
The variance estimator is updated to
\begin{equation} \label{eq:variance_gamma}
	\vartg{\ybb} = \frac{1}{\left( \tfrac{n}{m} - 1 \right) \left( \nb \right)}  \sum_{j=1}^{\nb} (\yb_j - \ybb)^2.
\end{equation}
With $\gamma=m$ and $n=mk$, (\ref{eq:variance_gamma}) reduces to (\ref{eq:var}).
With $\gamma=1$ (the maximally-overlapping case), the denominator in (\ref{eq:variance_gamma}) is slightly different than the one in  \citep{Meketon1984}.
This slightly different denominator 
makes $\vto{\ybb}$ an unbiased estimator for i.i.d.\ data for all finite $m$ and $n$ with the same asymptotic benefits \citep{Song1992}.

The amount of asymptotic reduction in $\var{\vartg{\ybb}}$ for a given $0\leq \gamma< m$, relative its nonoverlapping counterpart 
%$\var{\vtm{\ybb}}$ 
depends on the asymptotic ratio of $\gamma/m$, which we denote by $\gammab$. 
The maximally-overlapping case ($\gamma = 1$) corresponds to $\gammab=0$, and the nonoverlapping case corresponds to $\gammab=1$. 
\cite{damerdji1994strong,damerdji1995mean} studied batching structures that guarantee almost sure and mean-square consistency of $\vartg{\ybb}$ in the nonoverlapping and maximally-overlapping cases. 
Under appropriate conditions, these results, along with those of \cite{Meketon1984} and \cite{Welch1987}, establish that 
\begin{equation}
\label{eq:varvarred}
\frac{n}{m}\var{n \vartg{\ybb}} \rightarrow \beta_{\gammab} \sigma^4
\end{equation}
as the batch size $m$ and the number of batches $n/m$ tend to infinity. 
In (\ref{eq:varvarred}), $\beta_1 = 2$ for nonoverlapping batches, $\beta_0 = \frac{4}{3}$ for maximally overlapping batches, and more generally, $\beta_{\gammab}$ is twice the value of 
\begin{equation} \label{eq:var_reduct_formula}
	\gammab \left( 1 + 2 \sum_{j=1}^{\cogb-1} (1-j\gammab)^2 \right).
\end{equation}
If $\ogb$ is an integer, (\ref{eq:var_reduct_formula}) simplifies to $\frac{2+\gammab^2}{3}$.
This means that the maximally-overlapping variance estimator has a variance that is two-thirds ($66.67\%$) of the nonoverlapping estimator.
As the amount of overlap decreases, the variance increases. 
When only half of the observations overlap ($\gammab = 1/2$ or $\gamma=m/4$), for instance, the variance is 75\% of the nonoverlapping counterpart.

A $(1-\alpha)$-level approximate CI on the mean can be formed using overlapping batches by $\ybb \pm t_{d_{\gamma}(k-1),\alpha/2}$ $\sqrt{\vartg{\ybb}}$, where the degrees of freedom increase $d_{\gamma}$ is the multiplicative inverse of (\ref{eq:var_reduct_formula}) \citep{Welch1987}.  
If $\ogb$ is an integer, $d_{\gamma} = \frac{3}{2+\gammab^2}$.  For the maximally-overlapping batch means, the degrees of freedom is $d_{\gamma} = \frac{3}{2}$.

@@

Taking $n$ observations of the random variables $\xit$ and splitting them into $k$ {\it nonoverlapping} batches of size $m$, define 
\begin{align} 
	\gb_j  & = \frac{1}{m} \sum_{i = 1}^{m} f(\xh, \xit^{m(j-1)+i}) - \min_{\x \in X} \frac{1}{m} \sum_{i=1}^{m} f(\x, \xit^{m(j-1)+i}), \ \ \ \ j = 1,2, \dots, k \label{eq:mrp_gap}\\
         & = \frac{1}{m} \sum_{i = 1}^{m} \left[ f(\xh, \xit^{m(j-1)+i}) -  f(\xs_j, \xit^{m(j-1)+i}) \right], \ \ \ \ j = 1,2, \dots, k, \notag
\end{align}
where $\xs_j$ denotes an optimal solution to (\ref{eq:sto_prog_m}) of batch $j$ given on the right-hand side of (\ref{eq:mrp_gap}).    
Since we are assessing the quality of a given solution $\xh \in \X$, we suppress it from the notation $\gb_j$.
The optimality gap estimator $\gb_j$ is similar to the nonoverlapping $\yb_j$ with individual observations $y^{m(j-1)+i} = f(\xh,\xit^{m(j-1)+i}) - f(\xs_j,\xit^{m(j-1)+i})$, for $i=1,\ldots,m$ and $j=1,\ldots, k$.
Notice that in this case, $y^i$ not only depends on observation $i$ of batch $j$, $\xit^{m(j-1)+i}$, but all the observations in that batch $\xit^{m(j-1)+1},\xit^{m(j-1)+2},\ldots,\xit^{m(j-1)+m}$ because $\xs_j$ is the optimal solution to (\ref{eq:sto_prog_m}) given in (\ref{eq:mrp_gap}).   
As before, after $k$ estimates are obtained, the overall mean of these estimates $\gbb = \frac{1}{k} \sum_{j=1}^k \gb_j$ provides a point estimator of the optimality gap.  
The variance estimator is obtained as in (\ref{eq:var}) by $\varh{\gbb} = \frac{1}{k} \frac{1}{k-1} \sum_{j=1}^k (\gb_j - \gbb)^2$, which results in an approximate $(1-\alpha)$-level one-sided CI $\left[0, \gbb + t_{k-1,\alpha} \sqrt{\varh{\gbb}} \right]$.

@@

@@

The results on overlapping batch means like (\ref{eq:varvarred}) occur in the limit as $n, m, k=n/m \rightarrow \infty$.  
Let $n_l, m_l$, and $k_l$ be sequences of numbers satisfying these requirements, then the limits will be taken as $l \rightarrow \infty$.  
We assume a batching structure where $m_l = \lfloor n_l ^r \rfloor$ for some $0<r<1$, as is typical in simulation, and $n_l$ tends to infinity as $l \rightarrow \infty$.  
In addition, we may desire that the batch nonoverlap parameter change with $l$, to ensure that $\gamma_l / m_l = \gammab_l$ converges to the constant $\gammab$.  
Now we are ready to define the estimators for OMRP.
\begin{align}
	\gb_{lj} & = \frac{1}{m_l} \sum_{i=1}^{m_l} \left[ f(\xh,\xit^{\gamma_l(j-1)+i}) - f(\xs_j,\xit^{\gamma_l(j-1)+1}) \right],\ \ \ j = 1, 2, \dots, \nbl, \label{eq:gbar} \\
	\gbb_l & = \frac{1}{n_l} \sum_{i=1}^{n_l} \frac{1}{|B(i)|} \sum_{j \in B(i)} \left[ f(\xh,\xiti) - f(\xs_j,\xiti) \right], \label{eq:gbb} \\
	VG_l & = \dfrac{1}{\left( \tfrac{n_l}{m_l} - 1 \right) \left(\nbl\right)} \sum_{j=1}^{\nbl} (\gb_{lj} - \gbb_l)^2. \label{eq:vg}
\end{align}

@@

\subsection{Lower Variances}
\label{ssec:var_reduct}

To prove OMRP estimators have asymptotically lower variances compared to the nonoverlapping MRP estimator, we wish to show that 
$$
(\tfrac{n_l}{m_l}) \left( \var{n_l VG_l} - \var{n_l VD_l} \right) \rightarrow 0;
$$
see, e.g., (\ref{eq:varvarred}).
It suffices to show $\sqrt{{n_l}/{m_l}} \left( n_l VG_l - n_l VD_l \right) \xrightarrow{L^2} 0$, where $\xrightarrow{L^2}$ denotes mean-square convergence.

\begin{theorem} \label{thm:varvar_conv}
	Suppose assumptions A1--A3 hold, $\gammab = \frac{\gamma_l}{m_l}$ is constant for all $l$, and the batching structure ensures  $\frac{n_l}{m_l}\var{n_l VD_l} \rightarrow \beta_{\gammab} \sigma^4_{\xh}$ where $\beta_{\gammab}$ is the twice the value of (\ref{eq:var_reduct_formula}). 
        Then, $\frac{n_l}{m_l}\var{n_l VG_l} \rightarrow \beta_{\gammab} \sigma^4_{\xh}$ as $l \rightarrow \infty$ with the same $\beta_{\gammab}$. 
\end{theorem}

@@

Theorem \ref{thm:varvar_conv} gives conditions under which the same asymptotic lower variances are achieved for the OMRP optimality gap variance estimator relative to its nonoverlapping counterpart MRP. 
A hypothesis of Theorem \ref{thm:varvar_conv} is a batching structure that ensures convergence of $\frac{n_l}{m_l}\var{n_l VD_l}$ to the appropriate $\beta_{\gammab} \sigma^4_{\xh}$, ensuring the mean-square consistency of $n_l VD_l$. 
\citet{damerdji1995mean} shows that mean-square convergence is guaranteed by $\tfrac{1}{2} + \tfrac{1}{4+\epsilon} < r < 1$ for both nonoverlapping and maximally overlapping batches under the batching structure $m_l = \lfloor n_l^r \rfloor$ and $4+\epsilon$ finite moments assumption A2$^\prime$.  
With A2, the range of $r$ that ensures mean-square consistency can be found by setting $\epsilon=0$.

%We will empirically examine the variance reduction for small sample sizes in \S \ref{sec:comp}.

%We end our theoretical properties in the next section by examining the CIs formed using OMRP.
